A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2021-10-19 analysis/paper/paper.Rmd
Reducing raster resolution to 10 m (i.e. using DTM10 from hoydedata.no)

# Introduction
The close association between Norwegian Stone Age sites located in coastal regions and prehistoric shorelines was firmly established by the geologist W.C. Brøgger and his son, the archaeologist A.W. Brøgger, in their respective theses of 1905 (see also @maccurdy1906). One of the major implications of this finding is that relative sea-level change can be used for the dating of the sites. If most Stone Age sites were situated on or close to the contemporary shoreline, a reconstruction of the trajectory of shoreline displacement can be used to date the original occupation of the sites based on their elevation relative to the present day sea level. The method of shoreline dating has consequently been fundamental to Stone Age research in Norway ever since, and is used both independently and to compliment the dating of sites where other sources of temporal data such as typological indicators or radiometric dates are either limited or not available. Given the coarse and fuzzy resolution of established typological frameworks, the vast amount of surveyed sites that only contain generic lithic debitage that could hail from any part of the period, and as the conditions for the preservation of organic material is typically poor in Norway, dating by means of shoreline displacement is often the only and most precise and accurate method by which one can hope to date the sites. Dating by means of shoreline displacement is consequently fundamental to our understanding of the Norwegian Stone Age. This is both because shoreline dating is fundamental to the temporal framework on which our understanding of the period is based, but also because the method is only applicable so long as the societies in question have continuously settled on or close to the contemporary shore-line. Consequently, adherence or deviation from this pattern has major implications for the socio-economic foundations of the societies in question.

>>>>

# Introduction
The close association between Norwegian Stone Age sites located in coastal regions and prehistoric shorelines was firmly established by the geologist W.C. Brøgger and his son, the archaeologist A.W. Brøgger, in their respective theses of 1905 (see also @maccurdy1906). One of the major implications of this finding is that relative sea-level change can be used for the dating of the sites. This follows from the fact that if most Stone Age sites were situated on or close to the contemporary shoreline, a reconstruction of the trajectory of shoreline displacement can be used to date the original occupation of the sites based on their elevation relative to the present day sea level. The method of shoreline dating has consequently been fundamental to Stone Age research in Norway ever since, and is used both independently and to compliment the dating of sites where other sources of temporal data such as typological indicators or radiometric dates are either limited or not available. Given the coarse and fuzzy resolution of established typological frameworks, the vast amount of surveyed sites that only contain generic lithic debitage that could hail from any part of the period, and as the conditions for the preservation of organic material is typically poor in Norway, dating by means of shoreline displacement is often the only and most precise and accurate method by which one can hope to date the sites. Dating by means of shoreline displacement is consequently fundamental to our understanding of the Norwegian Stone Age. This is both because shoreline dating is fundamental to the temporal framework on which our understanding of the period is based, but also because the method is only applicable so long as the societies in question have continuously settled on or close to the contemporary shore-line. Consequently, adherence or deviation from this pattern has major implications for the socio-economic foundations of the societies in question.
